### 2020-05-11 initial/notes.tex
stash for princeton presentation

    \newcommand*{\scinot}[2]{#1\times10^{#2}}
    \newcommand*{\dotp}[2]{\left<#1\,\middle|\,#2\right>}
    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand*{\ptd}[2]{\frac{\partial^2 #1}{\partial#2^2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\pvec}[1]{\vec{#1}^{\,\prime}}
    \newcommand*{\svec}[1]{\vec{#1}\;\!}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \newcommand*{\ang}[0]{\;\text{\AA}}
    \newcommand*{\mum}[0]{\;\upmu \mathrm{m}}
    \newcommand*{\at}[1]{\left.#1\right|}

@@

    \DeclarePairedDelimiter\bra{\langle}{\rvert}
    \DeclarePairedDelimiter\ket{\lvert}{\rangle}
    \DeclarePairedDelimiter\abs{\lvert}{\rvert}
    \DeclarePairedDelimiter\ev{\langle}{\rangle}
    \DeclarePairedDelimiter\p{\lparen}{\rparen}
    \DeclarePairedDelimiter\s{\lbrack}{\rbrack}
    \DeclarePairedDelimiter\z{\lbrace}{\rbrace}

@@

This Hamiltonian is Kassandras Eq.\ 13, in the co-rotating frame with the
perturber's angular momentum:
\begin{equation}
    \mathcal{H} = -\frac{1}{2}\p*{\hat{s} \cdot \hat{l}}^2
        + \eta \p*{\hat{s} \cdot \hat{l}_p}.
\end{equation}
In this frame, we can choose $\hat{l} \equiv \hat{z}$ fixed, and $\hat{l}_p =
\cos I\hat{z} + \sin I\hat{x}$ fixed as well. Then
\begin{equation*}
    \hat{s} = \cos\theta \hat{z}
        - \sin\theta\p*{\sin \phi \hat{y} + \cos \phi \hat{x}}.
\end{equation*}
We can choose the convention for $\phi$ azimuthal angle requiring $\phi =
0, \pi$ mean coplanarity between $\hat{s}, \hat{l}, \hat{l}_p$ in the $\hat{x},
\hat{z}$ plane such that $\hat{l}_p, \hat{s}$ lie on the same side of
$\hat{l}$. Then we can evaluate in coordinates
\begin{align*}
    \hat{s} \cdot \hat{l} &= \cos \theta,\\
    \hat{s} \cdot \hat{l}_p
        &= \cos \theta \cos I - \sin I \sin \theta \cos \phi,\\
    \mathcal{H} &= -\frac{1}{2}\cos^2\theta
        + \eta \p*{\cos \theta \cos I - \sin I \sin \theta \cos \phi}.
\end{align*}
Note that if we take $\cos\theta$ to be our canonical variable, $\sin\theta =
\sqrt{1 - \cos^2\theta}$ can be used.

@@

The correct EOM comes from Kassandra's Eq.\ 12:
\begin{align*}
    \rd{\hat{s}}{t} &=
        \p*{\hat{s} \cdot \hat{l}}\p*{\hat{s} \times \hat{l}}
            -\eta\p*{\hat{s} \times \hat{l}_p},\\
        &= \p*{s_ys_z - \eta s_y\cos(I)}\hat{x}
            - \p*{s_xs_z + \eta \p*{s_x\cos I - s_z\sin I}}\hat{y}
            + \eta s_y \sin(I)\hat{z}.
\end{align*}

Alternatively, consider Hamilton's equations applied to the Hamiltonian:
\begin{align}
    \pd{\phi}{t} = \pd{\mathcal{H}}{(\cos\theta)}
        &= -\cos\theta + \eta\p*{\cos I + \sin I \cot \theta \cos \phi},\\
    \pd{(\cos \theta)}{t} = -\pd{\mathcal{H}}{\phi}
        &= -\eta \sin I \sin \theta \sin \phi.\label{eq:H_eom}
\end{align}
This produces the same trajectories as the Cartesian EOM, so this is correct.
However, since $\pd{\phi}{t} \propto 1/\sin\theta$, this is not a desirable
system of equations to use, as they are very stiff near $\theta \approx 0$.

@@

    \item For $\mu \approx 0$, we have instead $\pd{\phi}{t} = -\mu\p*{1 - \eta
        \sin I \cos \phi} + \eta \cos I = 0$. This forces $\mu_\pm = \frac{\eta
        \cos I}{1 \pm \eta \sin I}$, where $\phi_{\pm} = \pi, 0$ respectively.

        Note that $\phi = 0, \mu \approx 0$ is conventionally CS4. The
        linearization locally has form $\pd{\delta \phi}{t} = -\delta \mu\p*{1 -
        \eta \sin I}$ and $\pd{\delta \mu}{t} = -\eta \sin I \delta \phi$, so
        the eigenvalues are $\approx \mp \sqrt{\eta \sin I}$, and the two
        eigenvectors are $\p*{1, \pm \sqrt{\eta \sin I}}$.
\end{itemize}

@@

We notate $\mu = \cos\theta$; note that CS4 is $\mu_4 \approx \frac{\eta \cos
I}{1 - \eta \sin I} \approx \eta \cos I$. Setting the Hamiltonian equal to its
value at CS4 gives
\begin{align*}
    H_4 &\equiv H\p*{\mu_4, \phi_4}
        \approx -\frac{\mu_4^2}{2} + \eta \mu_4 \cos I - \eta \sin I,\\
        &= +\frac{\eta^2\cos^2 I}{2} - \eta \sin I,\\
    H(\mu_{sep}, \phi_{sep})
        &= H_4 = -\eta \sin I \cos \phi_{sep} - \frac{\mu_{sep}^2}{2} + \eta
            \mu_{sep} \cos I + \mathcal{O}(\eta^3),\\
    0 &\approx \frac{\mu_{sep}^2}{2} - \eta \mu_{sep} \cos I
        - \eta \sin I\p*{1 - \cos \phi_{sep}} + \frac{\eta^2\cos^2I}{2},\\
    \mu_{sep}\p*{\phi} &\approx \sqrt{2\eta \sin I\p*{1 - \cos \phi}}
        + \mathcal{O}(\eta).
\end{align*}

@@

We can add a tidal dissipation term; we write it in form
$\p*{\rd{\hat{s}}{t}}_{tide} = \epsilon \hat{s} \times \p*{\hat{l} \times
\hat{s}} = \epsilon\p*{\vec{l} - \p*{\vec{s} \cdot \vec{l}}\vec{s}}$. Expanding,
\begin{align}
    \p*{\rd{\hat{s}}{t}}_{tide} &= \epsilon \p*{\hat{z} - s_z\hat{s}}
        ,\nonumber\\
        &= \epsilon \p*{-s_z s_x\hat{x} - s_zs_y\hat{y} + \p*{1 - s_z^2}
            \hat{z}}.\label{eq:p1_eom}
\end{align}
We run numerical simulations for weaker $\epsilon \ll \eta \ll 1$ and stronger
$\epsilon \lesssim \eta \ll 1$.

We can seek equilibria of the the system including tides, which requires
\begin{align*}
    0 &= s_ys_z - \eta s_y\cos I - \epsilon s_zs_x,\\
    0 &= -s_xs_z - \eta\p*{s_x \cos I - s_z \sin I} - \epsilon s_zs_y,\\
    0 &= \eta s_y\sin(I) + \epsilon \p*{1 - s_z^2}.
\end{align*}
We expect at least two equilibria, based on the simulations: one near $s_z
\approx 1$ and one $s_z \approx 0$.

For near alignment/near Cassini state $1$, $1 - s_z \sim 1 - s_{\perp}^2$, so we
can set $s_z = 1$ to first order: $s_y - \epsilon s_x - \eta s_y \cos I = -s_x -
\eta\p*{s_x \cos I - \sin I} - \epsilon s_y = \eta s_y \sin I = 0$. This can be
satisfied if we set $s_x = \tan(I) \ll 1, s_y = \mathcal{O}\p*{\epsilon s_x}$;
this coarsely corresponds to Cassini state $1$.

The other solution should be near Cassini state $2$, where $s_x \approx 1$;
dropping second order terms forces $\eta s_y + \epsilon s_z = -s_z -
\eta\p*{\cos I - s_z\sin I} = \eta s_y \sin(I) + \epsilon = 0$. This can thus be
satisfied for $s_y \approx -\frac{\epsilon}{\eta \sin(I)}$. Thus, this explains
why as $\epsilon$ is increased, we first start to get points that don't converge
to Cassini state $2$ in the absence of tides, before starting to see points that
fail to converge to Cassini state $1$.

@@

We zoom in on Cassini State $4$, which has $\theta_4 =
-\frac{\pi}{2} + \frac{\eta \cos I}{1 - \eta \sin I}, \mu_4 = \frac{\eta \cos
I}{1 - \eta \sin I}, \phi_4 = 0$. Then, using
equations of motion
\begin{align}
    \pd{\phi}{t} &= -\mu + \eta\p*{\cos I + \sin I \frac{\mu}{\sqrt{1 - \mu^2}}
        \cos \phi} ,\\
    \pd{\mu}{t} &= -\eta \sin I \sin \phi + \s*{\epsilon \p*{1 - \mu^2}},
\end{align}
we can perturbatively require $\pd{\theta}{t} = 0$ for $\epsilon \neq 0$. This
corresponds to $\eta \sin I \sin \p*{\phi_4 + \delta \phi} \approx \epsilon$, or
$\delta \phi_4 = +\frac{\epsilon}{\eta \sin I}$. This is in agreement with
Dong's result.

@@

We notice that the separatrix is a heteroclinic orbit, or a saddle connection,
in the dissipation free problem. Introducing dissipation breaks the saddle
connection by a distance that can be estimated with the Melnikov distance. This
is G\&H Equation 4.5.11 or something:
\begin{align}
    d(t_0) &= \frac{\epsilon M(t_0)}{\abs*{f(q^0(0))}} + \mathcal{O}(\epsilon^2)
        ,\\
    M(t_0) &= \int\limits_{-\infty}^\infty
        \s*{f \times g}_{hetero}\;\mathrm{d}t.
\end{align}

@@

Let's apply this to the Cassini state Hamiltonian w/ dissipation. We first write
down our EOM in Melnikov form (we use canonical variables $\mu, \phi$):
\begin{align}
    \rd{\hat{s}}{t}
        &= \underbrace{
            \pd{\mathcal{H}}{\mu}\hat{\phi} - \pd{\mathcal{H}}{\phi}\hat{\mu}}_f
            + \epsilon \underbrace{\p*{1 - \mu^2}\hat{\mu}}_g.
\end{align}
Then $f \times g = f_\phi g_\mu = \pd{\mathcal{H}}{\mu}\p*{1 - \mu^2}$. We then
want to integrate this along the heteroclinic orbit. We can make change of
variables
\begin{equation}
    M = \int\limits_0^{2\pi}\pd{\mathcal{H}}{\mu}
        \p*{1 - \mu^2}\p*{\pd{\phi}{t}}^{-1}\;\mathrm{d}\phi.
\end{equation}
But thankfully, $\pd{\mathcal{H}}{\mu} = \pd{\phi}{t}$ in the absence of
dissipation, and so $M = 2\pi\p*{1 - \mu^2} \approx 2\pi$.
Thus, the Melnikov distance at point $q^0$, a point on the heteroclinic
orbit of the unperturbed Hamiltonian, is just
\begin{equation}
    d(q^0) = \frac{2\pi \epsilon}{\abs*{f(q^0)}}.
\end{equation}
Note that the maximum value $\abs*{f(q^0)}$, which occurs at $\phi = \pi$, is
just $f_{\max} \approx \sqrt{4\eta \sin I}$.

@@

My $\mu > \mu_{sep}$ simulations don't seem to work very well, so I'll focus on
the $\delta \mu^{-1}$ case. In this case, define $\delta \mu_c: \Delta
\mu\p*{\delta \mu_c} = -\delta \mu_c$, i.e.\ the point that jumps immediately to
the saddle point. Furthermore, assume the inbound distribution is flat between
$\delta \mu_c, f^{-1}\p*{\delta \mu_c}$. TODO\@: empirically, $\mu_c \sim
\epsilon T$ is \emph{flat} with $\eta$, probably just because we're not getting
sufficiently close to the saddle point for the $\propto \sqrt{\eta}$ to kick in.

Then, we can compare the empirical Poincar\'e section of the points that cross
the separatrix versus the total predicted interval width $\delta \mu_c,
f^{-1}\p*{\delta \mu_c}$; this would predict $7.2\%, 18\%$. This does alright!

@@

Now, let's reconsider the Melnikov integral when perturbing this finite
(non-homoclinic orbit); this may no longer be an exact result but should give
the correct scaling:
\begin{equation}
    M_c = \int\limits_0^T \pd{\phi}{t}\epsilon \p*{1 - \mu^2}\;\mathrm{d}t.
\end{equation}
Naively, we might claim that, since $\pd{\phi}{t}$ changes signs halfway through
the interval of integration, that the only surviving component is $2\epsilon
\bar{\mu}\mu'$, where $\bar{\mu} = \mu_4$ is the average value of $\mu$ and
$\mu'$ is the fluctuation. This gives
\begin{equation}
    M_c = 2\int\limits_0^{2\pi}2\epsilon \frac{\eta \cos I}{1 + \eta \sin I}
        \sqrt{2\eta \sin I \p*{1 - \cos \phi}}\;\mathrm{d}\phi.
\end{equation}
Note that $M_c \propto \eta^{3/2}$, and since the gap opened $\Delta_c =
\frac{M_c}{\pd{\phi}{t}} \propto \eta$, it seems like we're on the right track.
Specifically:
\begin{align}
    M_c &\approx \epsilon 2\eta \cos I A_{sep},\\
    \Delta_c &\approx 2\epsilon \eta \cos I \p*{16 \sqrt{\eta \sin I}}
        \frac{1}{\sqrt{4\eta \sin I}},\\
        &\approx 16\epsilon \eta \cos I.
\end{align}
This also agrees exceedingly well with our simulations This then gives us
hopping probability
\begin{equation}
    P_{hop} = \frac{\Delta_c}{\Delta_M} \approx
        \frac{16\eta^{3/2}\cos I \sqrt{\sin I}}{\pi}.
\end{equation}
This agrees perfectly with the cases we've run.

\textbf{Note:} The full formula, by actually evaluating all the terms, comes out
to be
\begin{equation}
    P_{hop} = \frac{16\eta^{3/2}\cos I \sqrt{\sin I}}{\pi
        \p*{1 - 2\eta \sin I - \eta^2 \cos^2 I}
        + 8\eta^{3/2}\cos I \sqrt{\sin I}}.
\end{equation}
It bears noting this formula slightly overpredicts $P_{hop}$ compared to the
actual simulation values; this comes down to ICs too near the separatrix being
necessarily escape-bound (can see \autoref{fig:CS4s} for more evidence).

@@

In order to probe the relevant physics here, we first cite without proof (refer
to \autoref{ss:revisited} for this) that $\pd{\Delta H}{\alpha} =
\frac{\mu^2}{2} - \frac{1}{2\alpha^2}\cos^2 I$. This implies that
\begin{equation}
    B_i = \int_{\mathcal{C}_i} \dot{\alpha}\s*{
        \frac{\mu^2}{2} - \frac{1}{2\alpha^2}\cos^2 I}
        \;\mathrm{d}t.
\end{equation}
Recall that the capture probability is $\frac{B_1 + B_2}{B_1}$ when this
expression is $\in [0, 1]$. However, we immediately see a problem: the first
term will dominate the second term, since $\alpha \gg 1$, but the first term is
also positive semidefinite. This implies $B_1, B_2$ will always have the same
sign unless $\dot{\alpha}$ has different signs over the two trajectories!

@@

Our expression then becomes:
\begin{align*}
    B_i &= \int -\frac{\epsilon \mu}{2\alpha^2}\s*{
        \alpha^2 \mu^2 - \cos^2 I}\;\mathrm{d}t,\\
        &\approx -\frac{\epsilon}{2\alpha^2} \int
            \mu \frac{\alpha^2 \mu^2 - \cos^2 I}{
                -\alpha \mu + \cos I}\;\mathrm{d}\phi,\\
        &\approx +\frac{\epsilon}{2\alpha^2}\int
            \mu\p*{\alpha \mu + \cos I}\;\mathrm{d}\phi.
\end{align*}
Now, we can evaluate our two integrals. For some reason my usual arguments
(taking symmetric/antisymmetric part etc.) don't seem to do too well here, there
are too many terms, so we will keep as many terms as possible.

First, we tackle the top integral. Here, the integrand is evaluated at
$\mu_+(\phi) = \eta \cos I + \sqrt{2\eta \sin I\p*{1 - \cos \phi}}$\footnote{
Bears noting that $\mu_{\pm} = \frac{\eta \cos I}{1 - \eta \sin I} \pm \dots$
significantly improves accuracy, it appears. Too many replacements to fix}, so
we can plug this in and compute
\begin{align*}
    B_{top}\frac{2\alpha^2}{\epsilon} &= \int\limits_{2\pi}^{0}
            \mu(\alpha \mu + \cos I)\;\mathrm{d}\phi,\\
        &= -\int\limits_0^{2\pi}
            \p*{\eta \cos I + \sqrt{2\eta \sin I\p*{1 - \cos \phi}}}
                \p*{2\cos I + \sqrt{2\sin I\p*{1 - \cos \phi}/\eta}}
                \;\mathrm{d}\phi.
\end{align*}
Let's track the four terms in FOIL (first, outer, inner, last) order. (F)
evaluates to just $4\pi \eta \cos^2 I$. (O) evaluates to $\frac{\cos
I}{2}A_{sep} = 8\cos I\sqrt{\eta \sin I}$, and (I) twice that. Lastly, (L)
evaluates to $4\pi \sin I$. This agrees with our numerical check.

>>>>

    \newcommand*{\scinot}[2]{#1\times10^{#2}}
    \newcommand*{\dotp}[2]{\left<#1\,\middle|\,#2\right>}
    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand*{\ptd}[2]{\frac{\partial^2 #1}{\partial#2^2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\pvec}[1]{\vec{#1}^{\,\prime}}
    \newcommand*{\svec}[1]{\vec{#1}\;\!}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \newcommand*{\ang}[0]{\;\text{\AA}}
    \newcommand*{\mum}[0]{\;\upmu \mathrm{m}}
    \newcommand*{\at}[1]{\left.#1\right|}
    \newcommand*{\bra}[1]{\left<#1\right|}
    \newcommand*{\ket}[1]{\left|#1\right>}
    \newcommand*{\abs}[1]{\left|#1\right|}
    \newcommand*{\ev}[1]{\langle#1\rangle}
    \newcommand*{\p}[1]{\left(#1\right)}
    \newcommand*{\s}[1]{\left[#1\right]}
    \newcommand*{\z}[1]{\left\{#1\right\}}

@@

@@

This Hamiltonian is Kassandras Eq.\ 13, in the co-rotating frame with the
perturber's angular momentum:
\begin{equation}
    \mathcal{H} = -\frac{1}{2}\p{\hat{s} \cdot \hat{l}}^2
        + \eta \p{\hat{s} \cdot \hat{l}_p}.
\end{equation}
In this frame, we can choose $\hat{l} \equiv \hat{z}$ fixed, and $\hat{l}_p =
\cos I\hat{z} + \sin I\hat{x}$ fixed as well. Then
\begin{equation*}
    \hat{s} = \cos\theta \hat{z}
        - \sin\theta\p{\sin \phi \hat{y} + \cos \phi \hat{x}}.
\end{equation*}
We can choose the convention for $\phi$ azimuthal angle requiring $\phi =
0, \pi$ mean coplanarity between $\hat{s}, \hat{l}, \hat{l}_p$ in the $\hat{x},
\hat{z}$ plane such that $\hat{l}_p, \hat{s}$ lie on the same side of
$\hat{l}$. Then we can evaluate in coordinates
\begin{align*}
    \hat{s} \cdot \hat{l} &= \cos \theta,\\
    \hat{s} \cdot \hat{l}_p
        &= \cos \theta \cos I - \sin I \sin \theta \cos \phi,\\
    \mathcal{H} &= -\frac{1}{2}\cos^2\theta
        + \eta \p{\cos \theta \cos I - \sin I \sin \theta \cos \phi}.
\end{align*}
Note that if we take $\cos\theta$ to be our canonical variable, $\sin\theta =
\sqrt{1 - \cos^2\theta}$ can be used.

@@

The correct EOM comes from Kassandra's Eq.\ 12:
\begin{align*}
    \rd{\hat{s}}{t} &=
        \p{\hat{s} \cdot \hat{l}}\p{\hat{s} \times \hat{l}}
            -\eta\p{\hat{s} \times \hat{l}_p},\\
        &= \p{s_ys_z - \eta s_y\cos(I)}\hat{x}
            - \p{s_xs_z + \eta \p{s_x\cos I - s_z\sin I}}\hat{y}
            + \eta s_y \sin(I)\hat{z}.
\end{align*}

Alternatively, consider Hamilton's equations applied to the Hamiltonian:
\begin{align}
    \pd{\phi}{t} = \pd{\mathcal{H}}{(\cos\theta)}
        &= -\cos\theta + \eta\p{\cos I + \sin I \cot \theta \cos \phi},\\
    \pd{(\cos \theta)}{t} = -\pd{\mathcal{H}}{\phi}
        &= -\eta \sin I \sin \theta \sin \phi.\label{eq:H_eom}
\end{align}
This produces the same trajectories as the Cartesian EOM, so this is correct.
However, since $\pd{\phi}{t} \propto 1/\sin\theta$, this is not a desirable
system of equations to use, as they are very stiff near $\theta \approx 0$.

@@

    \item For $\mu \approx 0$, we have instead $\pd{\phi}{t} = -\mu\p{1 - \eta
        \sin I \cos \phi} + \eta \cos I = 0$. This forces $\mu_\pm = \frac{\eta
        \cos I}{1 \pm \eta \sin I}$, where $\phi_{\pm} = \pi, 0$ respectively.

        Note that $\phi = 0, \mu \approx 0$ is conventionally CS4. The
        linearization locally has form $\pd{\delta \phi}{t} = -\delta \mu\p{1 -
        \eta \sin I}$ and $\pd{\delta \mu}{t} = -\eta \sin I \delta \phi$, so
        the eigenvalues are $\approx \mp \sqrt{\eta \sin I}$, and the two
        eigenvectors are $\p{1, \pm \sqrt{\eta \sin I}}$.
\end{itemize}

@@

We notate $\mu = \cos\theta$; note that CS4 is $\mu_4 \approx \frac{\eta \cos
I}{1 - \eta \sin I} \approx \eta \cos I$. Setting the Hamiltonian equal to its
value at CS4 gives
\begin{align*}
    H_4 &\equiv H\p{\mu_4, \phi_4}
        \approx -\frac{\mu_4^2}{2} + \eta \mu_4 \cos I - \eta \sin I,\\
        &= +\frac{\eta^2\cos^2 I}{2} - \eta \sin I,\\
    H(\mu_{sep}, \phi_{sep})
        &= H_4 = -\eta \sin I \cos \phi_{sep} - \frac{\mu_{sep}^2}{2} + \eta
            \mu_{sep} \cos I + \mathcal{O}(\eta^3),\\
    0 &\approx \frac{\mu_{sep}^2}{2} - \eta \mu_{sep} \cos I
        - \eta \sin I\p{1 - \cos \phi_{sep}} + \frac{\eta^2\cos^2I}{2},\\
    \mu_{sep}\p{\phi} &\approx \sqrt{2\eta \sin I\p{1 - \cos \phi}}
        + \mathcal{O}(\eta).
\end{align*}

@@

We can add a tidal dissipation term; we write it in form
$\p{\rd{\hat{s}}{t}}_{tide} = \epsilon \hat{s} \times \p{\hat{l} \times
\hat{s}} = \epsilon\p{\vec{l} - \p{\vec{s} \cdot \vec{l}}\vec{s}}$. Expanding,
\begin{align}
    \p{\rd{\hat{s}}{t}}_{tide} &= \epsilon \p{\hat{z} - s_z\hat{s}}
        ,\nonumber\\
        &= \epsilon \p{-s_z s_x\hat{x} - s_zs_y\hat{y} + \p{1 - s_z^2}
            \hat{z}}.\label{eq:p1_eom}
\end{align}
We run numerical simulations for weaker $\epsilon \ll \eta \ll 1$ and stronger
$\epsilon \lesssim \eta \ll 1$.

We can seek equilibria of the the system including tides, which requires
\begin{align*}
    0 &= s_ys_z - \eta s_y\cos I - \epsilon s_zs_x,\\
    0 &= -s_xs_z - \eta\p{s_x \cos I - s_z \sin I} - \epsilon s_zs_y,\\
    0 &= \eta s_y\sin(I) + \epsilon \p{1 - s_z^2}.
\end{align*}
We expect at least two equilibria, based on the simulations: one near $s_z
\approx 1$ and one $s_z \approx 0$.

For near alignment/near Cassini state $1$, $1 - s_z \sim 1 - s_{\perp}^2$, so we
can set $s_z = 1$ to first order: $s_y - \epsilon s_x - \eta s_y \cos I = -s_x -
\eta\p{s_x \cos I - \sin I} - \epsilon s_y = \eta s_y \sin I = 0$. This can be
satisfied if we set $s_x = \tan(I) \ll 1, s_y = \mathcal{O}\p{\epsilon s_x}$;
this coarsely corresponds to Cassini state $1$.

The other solution should be near Cassini state $2$, where $s_x \approx 1$;
dropping second order terms forces $\eta s_y + \epsilon s_z = -s_z -
\eta\p{\cos I - s_z\sin I} = \eta s_y \sin(I) + \epsilon = 0$. This can thus be
satisfied for $s_y \approx -\frac{\epsilon}{\eta \sin(I)}$. Thus, this explains
why as $\epsilon$ is increased, we first start to get points that don't converge
to Cassini state $2$ in the absence of tides, before starting to see points that
fail to converge to Cassini state $1$.

@@

We zoom in on Cassini State $4$, which has $\theta_4 =
-\frac{\pi}{2} + \frac{\eta \cos I}{1 - \eta \sin I}, \mu_4 = \frac{\eta \cos
I}{1 - \eta \sin I}, \phi_4 = 0$. Then, using
equations of motion
\begin{align}
    \pd{\phi}{t} &= -\mu + \eta\p{\cos I + \sin I \frac{\mu}{\sqrt{1 - \mu^2}}
        \cos \phi} ,\\
    \pd{\mu}{t} &= -\eta \sin I \sin \phi + \s{\epsilon \p{1 - \mu^2}},
\end{align}
we can perturbatively require $\pd{\theta}{t} = 0$ for $\epsilon \neq 0$. This
corresponds to $\eta \sin I \sin \p{\phi_4 + \delta \phi} \approx \epsilon$, or
$\delta \phi_4 = +\frac{\epsilon}{\eta \sin I}$. This is in agreement with
Dong's result.

@@

We notice that the separatrix is a heteroclinic orbit, or a saddle connection,
in the dissipation free problem. Introducing dissipation breaks the saddle
connection by a distance that can be estimated with the Melnikov distance. This
is G\&H Equation 4.5.11 or something:
\begin{align}
    d(t_0) &= \frac{\epsilon M(t_0)}{\abs{f(q^0(0))}} + \mathcal{O}(\epsilon^2)
        ,\\
    M(t_0) &= \int\limits_{-\infty}^\infty
        \s{f \times g}_{hetero}\;\mathrm{d}t.
\end{align}

@@

Let's apply this to the Cassini state Hamiltonian w/ dissipation. We first write
down our EOM in Melnikov form (we use canonical variables $\mu, \phi$):
\begin{align}
    \rd{\hat{s}}{t}
        &= \underbrace{
            \pd{\mathcal{H}}{\mu}\hat{\phi} - \pd{\mathcal{H}}{\phi}\hat{\mu}}_f
            + \epsilon \underbrace{\p{1 - \mu^2}\hat{\mu}}_g.
\end{align}
Then $f \times g = f_\phi g_\mu = \pd{\mathcal{H}}{\mu}\p{1 - \mu^2}$. We then
want to integrate this along the heteroclinic orbit. We can make change of
variables
\begin{equation}
    M = \int\limits_0^{2\pi}\pd{\mathcal{H}}{\mu}
        \p{1 - \mu^2}\p{\pd{\phi}{t}}^{-1}\;\mathrm{d}\phi.
\end{equation}
But thankfully, $\pd{\mathcal{H}}{\mu} = \pd{\phi}{t}$ in the absence of
dissipation, and so $M = 2\pi\p{1 - \mu^2} \approx 2\pi$.
Thus, the Melnikov distance at point $q^0$, a point on the heteroclinic
orbit of the unperturbed Hamiltonian, is just
\begin{equation}
    d(q^0) = \frac{2\pi \epsilon}{\abs{f(q^0)}}.
\end{equation}
Note that the maximum value $\abs{f(q^0)}$, which occurs at $\phi = \pi$, is
just $f_{\max} \approx \sqrt{4\eta \sin I}$.

@@

My $\mu > \mu_{sep}$ simulations don't seem to work very well, so I'll focus on
the $\delta \mu^{-1}$ case. In this case, define $\delta \mu_c: \Delta
\mu\p{\delta \mu_c} = -\delta \mu_c$, i.e.\ the point that jumps immediately to
the saddle point. Furthermore, assume the inbound distribution is flat between
$\delta \mu_c, f^{-1}\p{\delta \mu_c}$. TODO\@: empirically, $\mu_c \sim
\epsilon T$ is \emph{flat} with $\eta$, probably just because we're not getting
sufficiently close to the saddle point for the $\propto \sqrt{\eta}$ to kick in.

Then, we can compare the empirical Poincar\'e section of the points that cross
the separatrix versus the total predicted interval width $\delta \mu_c,
f^{-1}\p{\delta \mu_c}$; this would predict $7.2\%, 18\%$. This does alright!

@@

Now, let's reconsider the Melnikov integral when perturbing this finite
(non-homoclinic orbit); this may no longer be an exact result but should give
the correct scaling:
\begin{equation}
    M_c = \int\limits_0^T \pd{\phi}{t}\epsilon \p{1 - \mu^2}\;\mathrm{d}t.
\end{equation}
Naively, we might claim that, since $\pd{\phi}{t}$ changes signs halfway through
the interval of integration, that the only surviving component is $2\epsilon
\bar{\mu}\mu'$, where $\bar{\mu} = \mu_4$ is the average value of $\mu$ and
$\mu'$ is the fluctuation. This gives
\begin{equation}
    M_c = 2\int\limits_0^{2\pi}2\epsilon \frac{\eta \cos I}{1 + \eta \sin I}
        \sqrt{2\eta \sin I \p{1 - \cos \phi}}\;\mathrm{d}\phi.
\end{equation}
Note that $M_c \propto \eta^{3/2}$, and since the gap opened $\Delta_c =
\frac{M_c}{\pd{\phi}{t}} \propto \eta$, it seems like we're on the right track.
Specifically:
\begin{align}
    M_c &\approx \epsilon 2\eta \cos I A_{sep},\\
    \Delta_c &\approx 2\epsilon \eta \cos I \p{16 \sqrt{\eta \sin I}}
        \frac{1}{\sqrt{4\eta \sin I}},\\
        &\approx 16\epsilon \eta \cos I.
\end{align}
This also agrees exceedingly well with our simulations This then gives us
hopping probability
\begin{equation}
    P_{hop} = \frac{\Delta_c}{\Delta_M} \approx
        \frac{16\eta^{3/2}\cos I \sqrt{\sin I}}{\pi}.
\end{equation}
This agrees perfectly with the cases we've run.

\textbf{Note:} The full formula, by actually evaluating all the terms, comes out
to be
\begin{equation}
    P_{hop} = \frac{16\eta^{3/2}\cos I \sqrt{\sin I}}{\pi
        \p{1 - 2\eta \sin I - \eta^2 \cos^2 I}
        + 8\eta^{3/2}\cos I \sqrt{\sin I}}.
\end{equation}
It bears noting this formula slightly overpredicts $P_{hop}$ compared to the
actual simulation values; this comes down to ICs too near the separatrix being
necessarily escape-bound (can see \autoref{fig:CS4s} for more evidence).

@@

In order to probe the relevant physics here, we first cite without proof (refer
to \autoref{ss:revisited} for this) that $\pd{\Delta H}{\alpha} =
\frac{\mu^2}{2} - \frac{1}{2\alpha^2}\cos^2 I$. This implies that
\begin{equation}
    B_i = \int_{\mathcal{C}_i} \dot{\alpha}\s{
        \frac{\mu^2}{2} - \frac{1}{2\alpha^2}\cos^2 I}
        \;\mathrm{d}t.
\end{equation}
Recall that the capture probability is $\frac{B_1 + B_2}{B_1}$ when this
expression is $\in [0, 1]$. However, we immediately see a problem: the first
term will dominate the second term, since $\alpha \gg 1$, but the first term is
also positive semidefinite. This implies $B_1, B_2$ will always have the same
sign unless $\dot{\alpha}$ has different signs over the two trajectories!

@@

Our expression then becomes:
\begin{align*}
    B_i &= \int -\frac{\epsilon \mu}{2\alpha^2}\s{
        \alpha^2 \mu^2 - \cos^2 I}\;\mathrm{d}t,\\
        &\approx -\frac{\epsilon}{2\alpha^2} \int
            \mu \frac{\alpha^2 \mu^2 - \cos^2 I}{
                -\alpha \mu + \cos I}\;\mathrm{d}\phi,\\
        &\approx +\frac{\epsilon}{2\alpha^2}\int
            \mu\p{\alpha \mu + \cos I}\;\mathrm{d}\phi.
\end{align*}
Now, we can evaluate our two integrals. For some reason my usual arguments
(taking symmetric/antisymmetric part etc.) don't seem to do too well here, there
are too many terms, so we will keep as many terms as possible.

First, we tackle the top integral. Here, the integrand is evaluated at
$\mu_+(\phi) = \eta \cos I + \sqrt{2\eta \sin I\p{1 - \cos \phi}}$\footnote{
Bears noting that $\mu_{\pm} = \frac{\eta \cos I}{1 - \eta \sin I} \pm \dots$
significantly improves accuracy, it appears. Too many replacements to fix}, so
we can plug this in and compute
\begin{align*}
    B_{top}\frac{2\alpha^2}{\epsilon} &= \int\limits_{2\pi}^{0}
            \mu(\alpha \mu + \cos I)\;\mathrm{d}\phi,\\
        &= -\int\limits_0^{2\pi}
            \p{\eta \cos I + \sqrt{2\eta \sin I\p{1 - \cos \phi}}}
                \p{2\cos I + \sqrt{2\sin I\p{1 - \cos \phi}/\eta}}
                \;\mathrm{d}\phi.
\end{align*}
Let's track the four terms in FOIL (first, outer, inner, last) order. (F)
evaluates to just $4\pi \eta \cos^2 I$. (O) evaluates to $\frac{\cos
I}{2}A_{sep} = 8\cos I\sqrt{\eta \sin I}$, and (I) twice that. Lastly, (L)
evaluates to $4\pi \sin I$. This agrees with our numerical check.
